\documentclass[../../main.tex]{subfiles}
\begin{document}
\subsubsection{Update}
The operation $update$ will return a Collection $coll!$ which is the same as the input Collection $coll?$
except for at index $idx?$. The existing member $coll?_{idx?}$ is replaced by the provided Value $v?$ at $idx?$ in $coll!$
such that
$$idx? \mapsto v? \in coll! ~ \land ~ idx? \mapsto coll?_{idx?} \not \in coll!$$
which is equivalent to $remove ~\pipe ~append$
\begin{zed}
  update(coll?, v?, idx?) \equiv append(remove(coll?, idx?), v?, idx?)
\end{zed}
The functionality of $update$ is further explained in the following schema.
\begin{schema}{Update[Collection, V, \nat]}
  idx? : \nat \\
  coll?, coll! : Collection \\
  v? : V \\
  update~\_ : Collection \cross V \cross \nat \bij Collection
  \where
  1 = \#~idx? \\
  coll! = update(coll?, v?, idx?) @ \\
  \t1 let ~ ~ coll' == \{~ i : \nat ~| ~ i \in 0~..~idx?\} \extract coll? \\
  \t1 \ \ \ ~~~ coll'' == head(coll') ~\cat ~ v?\\
  \t1 \ \ \ ~~~ coll''' == \{~ j : \nat ~| ~ j \in idx?+1~..~\#~coll?\} \extract coll? \\
  \t1 =  coll'' ~\cat ~coll'' \implies \\
  \t2 (append(remove(coll', idx?), v?, idx?) \cat  coll'') ~ \land \\
  \t2 (~ v? \mapsto idx? \in coll!) ~\land \\
  \t2 (~\#~ coll! = \#~ coll?) ~\land
\end{schema}
The value which previously existed at $idx? \in coll?$ is replaced with $v?$ to result in $coll!$
\begin{itemize}
\item $coll'$ is the items in $coll?$ up to and including $idx?$
\item $coll''$ is the items in $coll?$ except the item at $idx?$ has been replaced with $v?$
\item $coll'''$ is the items in $coll?$ from $idx?+1$ to $\#~coll? \implies coll?_{idx?} \not \in coll''$
\end{itemize}
The following example illustrates these properties.
\begin{argue}
  X = \langle x_{0}, x_{1}, x_{2} \rangle \\
  \t1 x_{0} = 0 \\
  \t1 x_{1} = foo \\
  \t1 x_{2} = \langle a, b, c \rangle \\
  \t1 v? = bar \\
  update(X, v?, 0) = \langle bar, foo, \langle a, b, c \rangle \rangle \\
  update(X, v?, 1) = \langle 0, bar, \langle a, b, c \rangle \rangle \\
  update(X, v?, 2) = \langle 0, foo, bar \rangle \\
  update(X, v?, 3) = \langle 0, foo, \langle a, b, c \rangle, bar \rangle \\
  update(X, v?, 4) = append(X, v?, 3) = update(X, v?, 3) \iff 3 \not \in \dom X
\end{argue}
\end{document}
