%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

\documentclass{article}

%%% Package imports
\usepackage[utf8]{inputenc}
\usepackage[ruled]{algorithm2e}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{zed-csp}
\usepackage{breqn}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{dateplot}
\usepgfplotslibrary{polar}
\usetikzlibrary{pgfplots.dateplot}
\usetikzlibrary{pgfplots.patchplots}
\usetikzlibrary{patterns}

\usepackage{graphicx}
\graphicspath{{../../resources/plots/}}

\usepackage{floatrow}

\usepackage{calc}
%%% \makeatletter\amparswitchfalse\makeatother\
\DeclareMarginSet{hangleft}{\setfloatmargins
{\hskip-\marginparwidth\hskip-\marginparsep}{\hfil}}
\floatsetup[widefigure]{margins=hangleft}
%%% ^ Figure formatting within Appendex A

\lstset{literate = {-}{-}1} % get dashs to show up

\pgfplotsset{compat=1.15}

\SetKw{KwBy}{by}

\title{Data Analytics and Visualization Environment for xAPI and the Total Learning Architecture: DAVE Learning Analytics Algorithms}
\author{Yet Analytics}

\begin{document}

\begin{titlepage}
  \maketitle
\end{titlepage}

\section*{Introduction}
This report introduces the initial learning analytics algorithms,
\textbf{timeline of learner success},\textbf{which assessment
  questions are the most difficult} and \textbf{rate of completions}.
In doing so, it establishes a set of style
guidelines for the reporting of algorithms and associated visualization templates.
This document will be updated to include additional learning analytics
questions presented elsewhere within this github repo in addition to other learning
analytics algorithms which have yet to be defined. Updates may also
address refinement of these algorithms and this document should be
understood to be an example of algorithm presentation and not the
final state of any defined algorithm.

$\\\\$
The structure of this documents is as follows:
\begin{enumerate}
\item A formal specification for xAPI written in Z
  and referenced within the formal specifications of learning
  analytics algorithms
\item An algorithm definition which will consist of:
  \begin{enumerate}
  \item an introduction for the algorithm
  \item the structure of the ideal input data
  \item how to retrieve input data from an LRS
  \item the statement parameters which the algorithm will utilize
  \item notices regarding data collected during the 2018 pilot test of
    the TLA
  \item a summary of the algorithm
  \item the formal specification of the algorithm
  \item pseudocode representation of the algorithm
  \item JSONSchema for the output of the algorithm
  \item a description of the associated visualization
  \item a prototype of the visualization
  \item a collection of suggestions describing how the algorithm could be
    adapted to improve the quality of the visualization prototype
  \end{enumerate}
\end{enumerate}
$\\\\\\\\\\\\$ %%% only introduction on the first page



\section{Rate of Completions}

As learners engage in activities supported by a learning ecosystem, they will build
up a history of learning experiences. When the digital resources of that learning ecosystem
adhere to a framework dedicated to supporting and understanding the
learner, such as the Total Learning Architecture (TLA), it becomes
possible to retell their learning story through data and data visualization. One important aspect of
that story is the rate of completion\footnote{\label{defOfCompletion}
  Completion can be defined by the presence of the verb completed or by the presence of
  $\$.result.completion$ set equal to true. In this algorithm,
  completion is defined by the presence of the verb completed
  regardless of $\$.result.completion$. This decision affects how
  statements are retrieved and filtered. In the case where completion
  is defined by $\$.result.completion$, the query to the LRS would not
  include the verb paramater and there would need to be a filtering
  process which looks for the presence of $\$.result.completion$ =
  true} of the various digital resources within the learning
ecosystem.
\subsection{Ideal Statements}

In order to accurately portray the rates of completion, there
are a few base requirements of the data produced by a Learning Record
Provider (LRP). They are as follows:
\begin{itemize}
\item statements describing a learner completing an activity
  should\footnote{\label{verbIRICompletion} See footnote 4}
  use the verb $http://adlnet.gov/expapi/verbs/completed$
\item statements describing a learner completing an activity should
  report if the learner was successful or not via
  $\$.result.success$
\item statement describing a learner completing a scored activity
  should report the learners score via $\$.result.score.raw$,
  $\$.result.score.min$ and $\$.result.score.max$
\item activites must be uniquely and consistently identified across
  all statements
\item The time at which a learner completed a learning activity must be recorded
  \begin{itemize}
  \item The timestamp should contain an appropriate level of specificity.
  \item ie. Year, Month, Day, Hour, Minute, Second, Timezone
  \end{itemize}
\item statements describing a learner completing an activity should
  report the amount of time taken to complete the activity via $\$.result.duration$
\end{itemize}

\subsection{Input Data Retrieval}

How to query an LRS via a GET request to the Statements Resource via
curl. The following section contains the appropriate parameters with
example values as well as the curl command necessary for making the
request.\footnote{\label{refMoreLink2} See footnote
  1.}\footnote{\label{refnoZ2} See footnote
  2.}\footnote{\label{refallTime2} See footnote 3.}

\begin{lstlisting}[frame=single]
Verb = "verb=http://adlnet.gov/expapi/verbs/completed"

Since = "since=2018-07-20T12:08:47Z"

Until = "until=2018-07-21T12:08:47Z"

Base = "https://example.endpoint/statements?"

endpoint = Base + Verb + "&" + Since + "&" + Until

Auth = Hash generated from basic auth

S = curl -X GET -H "Authorization: Auth"
         -H "Content-Type: application/json"
         -H "X-Experience-API-Version: 1.0.3"
         Endpoint
\end{lstlisting}

\subsection{Statement Parameters to Utilize}

The statement parameter locations here are written in
\href{http://goessner.net/articles/JsonPath/}{JSONPath}. This notation
is also compatable with the xAPI Z notation due to the defined
hierarchy of components. Within the Z specifications, a variable name
will be used instead of the $\$$
\begin{itemize}
\item $\$.timestamp$
\item $\$.object.id$
\end{itemize}

\subsection{2018 Pilot TLA Statement Problems}

The initial pilot test data supports the core requirements of this
algorithm but completion statements only reports completion scores via
$\$.result.scaled$ instead of $\$.result.score.raw$,
$\$.result.score.min$ and
$\$.result.score.max$.\footnote{\label{scaledScores} The one potential
  issue with using scaled score is the
  calculation of scaled is not stricly defined by the xAPI
  specification but is instead up to the authors of the LRP. This
  results in the inability to reliably compare scaled scores across
  LRPs. if $\$.result.score.raw$ , $\$.result.score.min$ and
  $\$.result.score.max$ are reported for all questions, it becomes
  possible to reliably compare scores across LRPs by generating a
  scaled score in a consistent way.} Given that the offical 2018 pilot
test is scheduled to take place on July 27th, 2018, this section may
require updates pending future data review.

\subsection{Summary}

\begin{enumerate}
  \item Query an LRS via a GET request to the statemetns endpoint
    using the paramters verb, since and until.
  \item group statements by their $\$.object.id$
  \item select time range unit for use within rate calculation. Will
    default to day.
  \item determine the amount of time between the first and last instance of a $\$.object.id$ (in seconds) and divide it by the time unit. ie if the unit is minute, you would divide by 60.
  \item calculate the rate by dividing the count of a group (2) by the number of time units covered by the statements (4) so that the rate is the number of completions per activity per time unit.
\end{enumerate}

\subsection{Formal Specification}

\subsubsection{Basic Types}

$TIMEUNIT$ :== $\{second\} | \{minute\} | \{hour\} | \{day\} |
\{week\} | \{month\} | \{year\}$

\subsubsection{System State}

\begin{schema}{RateOfCompletion}
  Statements \\
  S_{completions} : \finset_1 \\
  S_{grouped},S_{timeunit}, S_{processed} : \finset \\
  \where
  S_{completions} = statements \\
  S_{grouped} = \{byId : \seq_1 statement\} \\
  S_{withRate} = \{byGroup: (\seq_1 statement, \nat)\} \\
  S_{processed} = \{rate : (id , \nat, TIMEUNIT)\}
\end{schema}
\begin{itemize}
\item The set $S_{completions}$ is a non-empty, finite set and is the
  component $statements$ which contains the results of the query to
  the LRS.
\item The sets $S_{grouped}$, $S_{withRate}$ and $S_{processed}$ are all finite sets
\item the set $S_{grouped}$ is a finite set of objects $byId$ which
  are non-empty, finite sequences of the component $statement$
\item the set $S_{withRate}$ is a finite set of objects $byGroup$ which
  are ordered pairs of non-empty, finite sequences of the component $statement$ and a natural number
\item the set $S_{processed}$ is a finite set of objects $rate$ where each
  contains the component $id$, a natural number and the type $TIMEUNIT$
\end{itemize}

\subsubsection{Initial System State}

\begin{schema}{InitRateOfCompletion}
  RateOfCompletion \\
  T : TIMEUNIT
  \where
  S_{completions} \not = \emptyset \\
  S_{grouped} = \emptyset \\
  S_{withRate} = \emptyset \\
  S_{processed} = \emptyset \\
  T = \{day\}
\end{schema}
\begin{itemize}
\item The set $S_{completions}$ is a non-empty set which contains the results of the GET request(s) to the LRS
\item The sets $S_{grouped}$ , $S_{withRate}$ and $S_{processed}$ are
  all initially empty
\item the variable T has the type $TIMEUNIT$ and the value $\{day\}$
\end{itemize}

\subsubsection{Calculate Rate}

\begin{schema}{IsoToUnix}
  convert : \finset_1 \fun \nat\#1 \\
  c? : \finset_1 \\
  c! : \nat\#1
  \where
  c! = convert(c?)
\end{schema}
\begin{itemize}
  \item The schema $IsoToUnix$ introduces the function $convert$ which
    takes in a finit set of one thing (a timestamp) and converts it to
    a single natural number.
  \item the purpose of this function is to convert an ISO 8601
    timestamp to the Unix epoch. The concrete definition of the conversion
    is outside the scope of this document
    \begin{itemize}
      \item The Unix epoch is the number of seconds that have elapsed
        since January 1, 1970 (midnight UTC/GMT), not counting leap seconds.
    \end{itemize}
  \end{itemize}

\begin{schema}{CalcRateByUnit}
  Statement \\
  IsoToUnix \\
  CountPerGroup \\
  unit? : TIMEUNIT \\
  s?,s! : \finset \\
  r : \nat \\
  rate : (\finset, TIMEUNIT) \fun \finset
  \where
  unit? \, = \,\{second\} \implies 1 \lor \{minute\} \implies 60
  \lor \{hour\} \implies 3600 \,\lor \\\t2 \{day\} \implies 86400 \lor
  \{week\} \implies 604800 \,\lor \\\t2 \{month\} \implies 2629743
  \lor \{year\} \implies 31556926 \\
  s? = \{g : seq_1 statement\} \\
  s! = rate(s?, unit?) \\
  s! = \{s : (g, r) \,|\, \forall g_{n}: g_{i}..g_{j} @ i \leq n \leq j @
  \exists \, s_{n} : (g_{n}, r_{n}) @ \\\t1
  r_{n} = count(g_{n}) \div ((convert(last~g_{n}.timestamp) - convert(head~g_{n}.timestamp)) \div unit?)\}
\end{schema}
\begin{itemize}
\item The schema $CalcRateByUnit$ introduces the function $rate$ where
  the input $s?$ is a set of objects $g$ which are each a non-empty,
  finite sequence of statements and the input $unit?$ represents a
  unit of time.
\item for every $g_{n}$ within the range $g_{i}..g_{j}$, there exists an associated object $s_{n}$ which is an orderd pair of $(g_{n}, r_{n})$ where $r_{n}$ is equal to the number of items within $g_{n}$ divided by the number of $unit?$s within the time range of $last~g_{n}.timestamp-head~g_{n}.timestamp$
\item the output of the function $rate$ is $s!$, the set of all $s_{n}$
\end{itemize}
$\\\\\\\\\\\\\\\\\\\\\\\\\\$ %%% keep header with z-schema
\subsubsection{Processes Results}
\begin{schema}{AggergateCompletionStatements}
  \Delta RateOfCompletion \\
  GroupByActivityId \\
  CalcRateByUnit \\
  grouped,processed,withRate : \finset \\
  r : \nat \\
  T? : TIMEUNIT
  \where
  T? = \{day\} \\
  grouped = \emptyset \\
  grouped' = group(S_{completions}) \\
  S_{grouped}' = S_{grouped} \cup grouped' \\
  withRate \subseteq S_{grouped}' \\
  withRate' = rate(withRate, T?) \\
  S_{withRate}' = withRate' \cup S_{withRate} \\
  processed \subseteq S_{withRate}' \\
  processed' = \{p: (id, r, T?) \,|\,
  \\\t3 \LET \{processed_{i}..processed_{j}\} == \{b_{i}..b_{j}\} @ \\\t3
  \forall b_{n} : b_{i}..b_{j} @ i \leq n \leq j @
  \exists \, p_{n} : (id_{n}, r_{n}, T?) @  \\\t3
  id_{n} = (head~(first~b_{n})).object.id \, \land \, \\\t3
  r_{n} = (second~b_{n})\} \\
  S_{processed}' = processed' \cup S_{processed}
\end{schema}

\begin{itemize}
\item The schema $AggergateCompletionStatements$ outlines how to calculate
  the rate of completion per $\$.object.id$ per $second|minute|hour|day|week|month|year$
  \begin{enumerate}
  \item $S_{grouped}'$ is the result of grouping the statements within $S_{completions}$ by their $\$.object.id$
  \item The groups from (1) are passed to the function $rate$ with the
    variable $T?$ which controls the unit of time, ie per day vs per week
  \item the result of (2) is then processed to create a triplet of
    $\$.object.id$, rate, unit of time for all unique $\$.object.id$
    within $S_{completions}$
\end{enumerate}
\end{itemize}

\subsubsection{Return}

\begin{schema}{ReturnAggergateCompletionStatements}
  \Xi RateOfCompletion \\
  AggergateCompletionStatements \\
  S_{processed}! : \finset
  \where
  S_{processed}! = S_{processed}
\end{schema}
\begin{itemize}
  \item The return value $S_{processed}!$ is equal to $S_{processed}$ after the operation described by $AggergateCompletionStatements$
\end{itemize}

\subsection{Pseudocode}

\begin{algorithm}[H]
  \SetAlgoLined
  \KwIn{$S_{completed}$, $timeUnit$}
  \KwResult{$ratePerObjTu'$}
  \emph{context = \{\}}\;
  \emph {ratePerObjTu = []}\;
  \While{$S_{completion} \not = \emptyset$}
  {\ForEach{$s \in S_{completion}$}
    {$id \leftarrow s.object.id$\;
      $ts \leftarrow convert(s.timestamp)$\;
      \eIf{$id \notin context$}
      {\bf do \\
        $times = [ts]$\;
        $context' \leftarrow \{id : times \}$\;
        $S_{completion}' \leftarrow S_{completion} \setminus s$\;
        recur $context', S_{completion}'$\;}
      {\bf do \\
        $times' \leftarrow context.id \cat ts$\;
        $context' \leftarrow \{id : times'\}$\;
        $S_{completion}' \leftarrow S_{completion} \setminus s$\;
        recur $context', S_{completion}'$\;}}}
  \ForEach {$k \in context'$}
  {$allTs \leftarrow context'.k$\;
    $totalDuration \leftarrow max(allTs) - min(allTs)$\;
    $totalCount \leftarrow count(allTs)$\;
    $rate \leftarrow totalCount \div (totalDuration \div timeUnit)$\;
    $subVec = [k, rate, timeUnit]$\;
    $ratePerObjTu' \leftarrow ratePerObjTu \cat subVec$\;
    {\bf recur} $ratePerObjTu'$\;}
 \Return $ratePerObjTu'$
  \caption{Rate of Completions}
\end{algorithm}
\begin{itemize}
\item Values from Z schemas are used within this pseudocode
\item the result of the algorithm is an array of arrays where each
  subarray contains a $statement.object.id$, the $rate$ and the
  $timeUnit$ used to calculate $rate$.
\end{itemize}
$\\$ %% header with text
\subsection{JSON Schema}

\begin{lstlisting}[]
{"type":"array",
   "items":{"type":"array",
      "items":[{"type":"string"}, {"type":"number"},
{"type":"string"}]}}
\end{lstlisting}

\subsection{Visualization Description}

The \textbf{Rate of Completions} visualization will be
a bar chart where the domain consists of $statement.object.id$ and the
range is a number greater than 0 (the rate of completions for that
$statement.object.id$). Every subarray within the array $ratePerObjTu$
will be a grouping within the bar chart. The pseudocode specifies an
input paramter $timeUnit$ which controls the calculation of the rate
(range of the visualization). $timeUnit$ could be per minute, per day,
per week, etc.

\subsection{Visualization prototype}

\pgfplotstabletypeset[string type]

\begin{tikzpicture}
  \begin{axis}[
    title = Rate of Completions,
    ylabel = Completions per Time Unit,
    symbolic x coords={question1-id,question2-id,question3-id},
    xtick=data]
    \addplot[ybar,fill=blue] coordinates{
      (question1-id,18)
      (question2-id,9)
      (question3-id,13)
    };
  \end{axis}
\end{tikzpicture}

\subsection{Prototype Improvement Suggestions}
Additional features may be implemented on top of this base
specification but they would require adding aditional values to each
subarray returned by the algorithm. These additional values can be
retrieved via (1) performing metadata lookup within or independently of the
algorithm (2) by utilizing additional xAPI statement paramters and/or (3) by
performing additional computations. The following examples assume the
metadata is contained within each statement available to the algorithm.

\begin{itemize}
\item use $statement.object.definition.name$ instead of
  $statement.object.id$ for x axis label
\item populate a tooltip with the people who have completed the
  activity. This could also include the number of times they have
  completed it.
\item populate a tooltip with the breakdown of which devices or platforms the
  activity was completed on. This would require the device type or platform to be
  reported within $statement.context.platform$
\item populate a tooltip with the breakdown of percentage successful
  for all completions of the activty. This would require
  $statement.result.success$
\item populate a tooltip with the breakdown of scores earned (if
  appliciable) for the completions. This would require
  $statement.result.score.raw$, $statement.result.score.min$ and
  $statement.result.score.max$
\item populate a tooltip with the competency assocaited with the
  completed activities. The competency should be reported
  via $statement.context.contextActivities$
\item populate a tooltip with the average duration spent to reach
  completions. This would require $statement.result.duration$ to be reported.
\end{itemize}
$\\$

\section{How Often are Recommendations Followed}
As learners engage in activities supported by a learning ecosystem, they will build
up a history of learning experiences. When the digital resources of that learning ecosystem
adhere to a framework dedicated to supporting and understanding the
learner, such as the Total Learning Architecture (TLA), it becomes
possible to retell their learning story through data and data
visualization. One important aspect of that story is the
recommendations provided to the learner and whether or not the learner
follows those recommendations.

\subsection{Ideal Statements}
In order to accurately determine if a learner is following
recommendations, there are a few requirements of the data produced by
a LRP and the recommender itself. They are as follows:
\begin{itemize}
\item Every time the recommender makes a recommendation, a statement
  should be produced which uses the verb
  $https://w3id.org/xapi/dod-isd/verbs/recommended$\footnote{\label{RecommendedIRI}
  See footnote 4} and has the
  recommended piece of content as the object.
  \begin{itemize}
    \item the content should be uniquely and consistently identified
      across all statements.
    \end{itemize}
\item When a learner launches recommended content, the resulting
  launched statement should use the verb
  $http://adlnet.gov/expapi/verbs/launched$\footnote{\label{LaunchedIRI}
  See footnote 4} and contain a refrence to the recommened content statement within
$\$.context.statement$
  \begin{itemize}
    \item Launching of content should use the above IRI regardless of
      why the content was launched
    \item If it not possible to refrence the exact recommended content
      statement, the launch statement should have some indication that
      it was the result of a recommendation.\footnote{\label{Recommendations} It is
    possible to determine if recommendations are followed (with some
    level of error) without this explicit linking of launched to
    recommended but this severly complicates the algorithm. In this
    case, in order to optmize for accuracy, the algorithm would need
    to consider the actor and their general activity within a session,
    the object of both launched and recommended statements generated
    within the session, the time lapse between recommendations and
    launches with a predefined lapse value which determines if a
    launch was close enough in time to a recommendation to be
    considered a result of the recommendation. An additonal constraint
    on the above case is the recommendation statements should contain
    a reference to to the person recieving the recommendation,
    otherwise determining the 1:1 relationships between recommendations and
    launches requires aditional complexity and will still not be 100\%
    accurate due to the reliance on the time lapse value.}
\end{itemize}
\end{itemize}

\subsection{Input Data Retrieval}
How to query an LRS via a GET request to the Statements Resource via
curl.\footnote{\label{refMoreLink2} footnote 1 applies to both S1
  and S2.}\footnote{\label{refnoZ2} See footnote
  2.}\footnote{\label{refallTime2} See footnote 3.}

\begin{lstlisting}[frame=single]
R = "verb=https://w3id.org/xapi/dod-isd/verbs/recommended"
L = "verb=http://adlnet.gov/expapi/verbs/launched"

Since = "since=2018-07-20T12:08:47Z"
Until = "until=2018-07-21T12:08:47Z"
Base = "https://example.endpoint/statements?"

endpoint1 = Base + R + "&" + Since + "&" + Until
endpoint2 = Base + L + "&" + Since + "&" + Until

Auth = Hash generated from basic auth

SR = curl -X GET -H "Authorization: Auth"
         -H "Content-Type: application/json"
         -H "X-Experience-API-Version: 1.0.3"
         endpoint1
SL = curl -X GET -H "Authorization: Auth"
         -H "Content-Type: application/json"
         -H "X-Experience-API-Version: 1.0.3"
         endpoint2

S = SR + SL
\end{lstlisting}

\subsection{Statement Parameters to Utilize}
The statement parameter locations here are written in
\href{http://goessner.net/articles/JsonPath/}{JSONPath}. This notation
is also compatable with the xAPI Z notation due to the defined
hierarchy of components. Within the Z specifications, a variable name
will be used instead of the $\$$
\begin{itemize}
\item $\$.verb.id$
\item $\$.context.statement$
\end{itemize}

\subsection{2018 Pilot TLA Statement Problems}
At the time of writing this document, launched statements do not
include a statement reference or any indication of a connection
between recommendations and launches. The authors of this document do
not have access to the LRS containing the recommended statements and
thus can not draw any conclusions about any issues which may be
present within those statements or any aspects of those statements
which may correlate them to launch statements. The following algorithm
is going to assume that the input set of statements follow the
guidlines outlined in section 5.1 as the additional algorithmic
considerations brought on by non ideal statements, as specified within
footnote 16, result in an algoirthm which is not optimal for near real
time visualizations.

\subsection{Summary}
\begin{enumerate}
  \item Query an LRS via a GET request to the statements endpoint
    using the paramters verb, since and until to gather all statements
    with the verb $http://adlnet.gov/expapi/verbs/launched$.
  \item Query an LRS via a GET request to the statements endpoint
    using the paramters verb, since and until to gather all statements
    with the verb $https://w3id.org/xapi/dod-isd/verbs/recommended$.\footnote{\label{sameSession} If since and until are specified,
      they should be the same in both requests.}
  \item Group all collections of statements by a $TIMEUNIT$
  \item seperate the collection of grouped launched statements into a
    collection of those which were the result of a recommendation and
    those which were not.
  \item Take the count of all groups of statements
    \begin{itemize}
    \item Recommended statements per $TIMEUNIT$
    \item Launches due to recommendations per $TIMEUNIT$
    \item Launches not due to recommendations per $TIMEUNIT$
    \end{itemize}
  \item Calculate summary statistics for the overall time range and
    per $TIMEUNIT$
    \begin{itemize}
    \item Divide launches due to recommendations by the total number of
      launches to determine the percentage of launches due to
      recommendations
    \item Divide launches due to recommendations by the total number
      of recommendations to determine the percentage of
      recommendations which are followed.
    \end{itemize}
\end{enumerate}

\subsection{Formal Specification}

\subsubsection{System State}

\begin{schema}{FollowedRecommendations}
  Statements \\
  CountPerGroup \\
  S_{recommended},S_{launched} : \finset_1 \\
  ordered_{L}, ordered_{R}, grouped_{launched}, grouped_{recommended}, \\
  onlyRecommended, cPerGroup_{launched}, cPerGroup_{recommended}, \\
  cPerGroup_{followed}, combined : \seq \\
  t_{start}, N_{launched}, N_{recommended}, N_{followed}, P_{followed}, P_{dueto}  : \nat \\
  tr_{start}, tr_{end} : \finset \\
  unit? : TIMEUNIT
  \where
  S_{recommended} = statements \\
  S_{launched} = statements \\
  combined = \langle (tr_{start}, tr_{end}, N_{launched}, N_{recommended},
  N_{followed}, P_{followed}, P_{dueto})\rangle \\
  count(grouped_{launched}) = count(grouped_{recommended}) \\
  count(onlyRecommended) = count(grouped_{launched}) \implies \\
  count(onlyRecommended) = count(grouped_{recommended}) \\
  count(cPerGroup_{launched}) = count(cPerGroup_{followed}) = count(cPerGroup_{recommended})
\end{schema}

\begin{itemize}
\item $S_{recommended}, S_{launched}$ are both non-empty, finite sets.
  \begin{itemize}
  \item $S_{recommended}$ and $S_{launched}$ contain
    the results of querying an LRS for recommended and launched
    statements respectively.
  \end{itemize}
\item $ordered_{L}, ordered_{R}, grouped_{launched},
  grouped_{recommended},$ $onlyRecommended, \\ cPerGroup_{launched},
  cPerGroup_{recommended}, cPerGroup_{followed}$ and $combined$ are
  all finite sequences.
  \begin{itemize}
    \item $ordered_{L}$ and $ordered_{R}$ are the sequences of
      statements within $S_{launched}$ and $S_{recommended}$
      respectively and sorted by timestamp.
    \item $grouped_{launched}$ is the result of grouping the
      statements within $ordered_{L}$ by $unit?$.
    \item $grouped_{recommended}$ is the result of grouping the
      statements within $ordered_{R}$ by $unit?$.
    \item $onlyRecommended$ is the result of filtering the statements
      within the sequence $grouped_{launched}$ to only include
      statements where $statement.context.statement$ is present
    \item $cPerGroup_{launched}, cPerGroup_{recommended},
      cPerGroup_{followed}$ are all sequences of numbers which
      represent the count within each subsequence of
      $grouped_{launched}$, $grouped_{recommended}$ and
      $onlyRecommended$ respectively.
    \item $combined$ is a sequence of ordered pairs where each pair
      consists of $tr_{start}$, $tr_{end}$, $N_{launched}$,
      $N_{recommended}$, $N_{followed}$, $P_{followed}$ and $P_{dueto}$
    \end{itemize}
\item $t_{start}, N_{launched}, N_{recommended}, N_{followed},
  P_{followed}, P_{dueto}$ are all natural numbers
\item $tr_{start}, tr_{end}$ are both timestamps which represent the
  the start and end of the time range for each a group of statements.
\item $unit?$ is an input representing a time interval, ie day vs
  month vs hour.
\item all sequences are the same length so that each subsequence
  represents the same time grouping. In other words, indexes are
  comparable across sequences.
\end{itemize}

\subsubsection{Initial System State}

\begin{schema}{InitFollowedRecommendations}
  FollowedRecommendations \\
  \where
  S_{recommended} \not = \emptyset \\
  S_{launched} \not = \emptyset \\
  unit? = \{day\} \\
  ordered_{L} = \langle  \rangle \\
  ordered_{R} = \langle  \rangle \\
  grouped_{launched} = \langle  \rangle \\
  grouped_{recommended} = \langle  \rangle \\
  onlyRecommended = \langle  \rangle \\
  cPerGroup_{launched} = \langle  \rangle \\
  cPerGropu_{recommended} = \langle  \rangle \\
  cPerGroup_{followed} = \langle  \rangle \\
  combined = \langle  \rangle \\
  t_{start} = 0 \\
  N_{launched} = 0 \\
  N_{recommended} = 0 \\
  N_{followed} = 0 \\
  P_{followed} = 0 \\
  P_{dueto} = 0 \\
\end{schema}

\begin{itemize}
  \item $S_{recommended}$ and $S_{launched}$ are initially not empty
    sets
  \item all sequences are initially empty
  \item all numbers are initially zero
  \item the default $TIMEUNIT$ is set to day
\end{itemize}

\subsubsection{Group by Timestamp}

\begin{schema}{SortByTimestamp}
  Statement \\
  IsoToUnix \\
  orderByTimestamp : \finset_1 \fun seq_1 \\
  o? : \finset_1 \\
  o! : seq_1 statement
  \where
  o? = \{o : statement\} \\
  o! = orderByTimestamp(o?) \\
  o! = \langle o_{i}..o_{j} \rangle @ \forall o_{n} : o_{i}..o_{j} @
  o_{n} : STATEMENT \land i \leq n \leq j @ \\\t3
  convert(o_{i}.timestamp) \leq convert(o_{n}.timestamp) \leq
  convert(o_{j}.timestamp)
\end{schema}

\begin{itemize}
\item The schema $SortByTimestamp$ introduces the function
  $orderByTimestamp$ which takes in a non-empty, finite set and
  returns a non-empty, finite sequence.
\item $orderByTimestamp$ is a sequence of statements ordered from
  earliest to latest.
\end{itemize}

\begin{schema}{WithinRange}
  withinRange : (\nat, \nat, \nat, TIMEUNIT) \fun \finset_1\#1 \\
  in?, start?, state? : \nat \\
  unit? : TIMEUNIT \\
  out! : \{TRUE\} \lor \{FALSE\}
  \where
  unit? \, = \,\{second\} \implies 1 \lor \{minute\} \implies 60
  \lor \{hour\} \implies 3600 \,\lor \\\t2 \{day\} \implies 86400 \lor
  \{week\} \implies 604800 \,\lor \\\t2 \{month\} \implies 2629743
  \lor \{year\} \implies 31556926 \\

  out! = withinRange(in?, start?, state?, unit?) \\
  withinRange(in?, start?, state?, unit?) = \\\t3 \IF in? \leq
  start? + ((state? + 1) * unit?) \\\t4 \THEN out! = \{TRUE\} \\\t4 \ELSE
  out! = \{FALSE\} \\
\end{schema}

\begin{itemize}
\item The schema $WithinRange$ introduces the function $withinRange$
  which takes in three numbers and a $TIMEUNIT$ and returns either
  $\{TRUE\}$ or $\{FALSE\}$
\item $withinRange$ checks to see if $in?$ is less than or equal to a
  start time $start?$ plus the result of multiplying the numeric
  conversion for $unit?$ by the $state?$.
\item $state?$ represents the current group, ie. day 1 vs day 2 vs day
  3. The $+ 1$ is to account for array indexes starting at 0.
\end{itemize}

\begin{schema}{GroupByTimeUnit}
  Statement \\
  IsoToUnix \\
  WithinRange \\
  groupByTimeUnit : (\seq_1, \nat, TIMEUNIT) \fun \seq_1 \\
  g?,g! : \seq_1 \\
  t_{start}? : \nat
  \where
  g? = \langle g?_{i}..g?_{j} \rangle @ \forall g?_{n} :
  g?_{i}..g?_{j} @ i \leq n \leq j @ g?_{n} = statement \,\land \\\t1
  convert(g?_{i}.timestamp) \leq convert(g?_{n}.timestamp) \leq
  convert(g?_{j}.timestamp) \\
  g! = groupByTimeUnit(g?, t_{start}? state?, unit?) \\
  g! = \langle g : seq \,|\, \forall g?_{n} :
  g?_{i}..g?_{j} @
  \exists_1 \langle g_{r} \rangle : \langle g_{q} \rangle
  .. \langle g_{s} \rangle @ q \leq r \leq s \land r = state? @
  \\\t3 \IF withinRange(convert(g?_{n}.timestamp), t_{start}, r, unit?) =
  \{TRUE\} \\\t4 \THEN g? \extract g?_{n} \,\land\, g?_{n} \inseq
  \langle g_{r} \rangle \implies \langle g_{r} \rangle = \langle
  g_{ri}..g_{rn}..g_{rj} \rangle @ ri \leq rn \leq rj @ g_{rn} =
  g?_{n}  \\\t4 \ELSE \IF \forall g_{n}? : g_{i}?..g_{j}? @
  withinRange(g?_{n}, t_{start}, r, unit?) = \{FALSE\} \\\t5 \THEN
  groupByTimeUnit((g? \extract \langle g_{r} \rangle) , t_{start}? (r
  + 1), unit?) @ \langle g_{r} \rangle = \langle  \rangle \, \lor \not
  = \langle  \rangle \rangle
\end{schema}

\begin{itemize}
\item The schema $GroupByTimeUnit$ intorudces the function
  $groupByTimeUnit$ which takes as arguments a non-empty, finite
  sequence, a natural number and a $TIMEUNIT$ and outputs a non-empty,
  finite sequence of sequences.
\item For every statement within the input sequence, $groupByTimeUnit$
  checks to see if the timestamp of that statement is within the range
  of $t_{start}$ and $unit?$. If it is, that statement is removed from
  the input sequence $g?$ and added to the current subsequence
  $\langle g_{r} \rangle$. If none of the remaining statements within
  the input sequence are within the range of $t_{start}$ and $unit?$,
  then the variable $state?$ is incremented, the current subsequence
  $\langle g_{r} \rangle$ is either a collection of matched statements or is an
  empty sequence and the search for remaining subsequences $\langle
  g_{r+state?} \rangle$ continues.
\item because the input sequence $g?$ is orderd chronologically, this
  implies that once a statement does not fit into a range, the rest of
  the statements remaining in the input sequence will not fit into
  that range and $state?$ must be incremented to generate a new
  subsequence $\langle g_{r + state?} \rangle$ so that the remaining
  statements can be grouped.
\end{itemize}
$\\$ %header with text
\subsubsection{Processes Results}

\begin{schema}{OrderStatements}
  \Delta FollowedRecommendations \\
  SortByTimestamp
  \where
  ordered_{L}' = orderByTimestamp(S_{launched}) \\
  ordered_{R}' = orderByTimestamp(S_{recommended}) \\
  t_{start}' = convert((head~ordered_{L}').timestamp) \\
\end{schema}

\begin{itemize}
  \item The schema $OrderStatements$ updates the system state defined
    by the schema $FollowedRecommendations$.
  \item $ordered_{L}'$ is the result of ordering the statements contained
    within the set $S_{launched}$ chronologically.
  \item $ordered_{R}'$ is the result of ordering the statements contained
    within the set $S_{recommended}$ chronologically.
  \item $t_{start}'$ is the timestamp from the first statement within
    $ordered_{L}'$ converted to unix time.
\end{itemize}

\begin{schema}{GroupByTime}
  \Delta FollowedRecommendations \\
  GroupByTimeUnit \\
  \where
  grouped_{launched}' = groupByTimeUnit(ordered_{L}', t_{start}', 0, unit?) \\
  grouped_{recommended}' = groupByTimeUnit(ordered_{R}', t_{start}',
  0, unit?)
\end{schema}

\begin{itemize}
  \item The schema $GroupByTime$ updates the state defined by the schema
    $FollowedRecommendations$.
  \item $grouped_{launched}'$ is the result of passing $ordered_{L}'$,
    $t_{start}'$, 0 and $unit?$ to the function $groupByTimeUnit$.
  \item $grouped_{recommended}'$ is the result of passing $ordered_{R}'$,
    $t_{start}'$, 0 and $unit?$ to the function $groupByTimeUnit$.
\end{itemize}

\begin{schema}{OnlyRecommendedLaunches}
  \Delta FollowedRecommendations
  \where
  onlyRecommended' = \langle o : seq \,|\, \LET grouped_{launched}'==gl
  \implies \\\t5 \langle \langle gl_{i} \rangle..\langle gl_{j}
  \rangle \rangle \implies \langle \langle gl_{ii}..gl_{ij}
  \rangle..\langle gl_{ji}..gl_{jj}\rangle\rangle @ \\\t5
  \forall \langle gl_{n} \rangle : \langle gl_{i} \rangle..\langle
  gl_{j} \rangle @ \exists_1 \langle o_{n} \rangle : \langle o_{i}
  \rangle..\langle o_{j} \rangle \implies \langle \langle o_{ii}..o_{ij} \rangle..\langle
  o_{ji}..o_{jj}\rangle\rangle @ \\\t5 ((\forall o_{in} : o_{ii}..o_{ij} @
  o_{in}.context.statement \not = \emptyset \land o_{in} \inseq
  gl_{i}) \land \\\t5 (\forall o_{jn} : o_{ji}..o_{jj} @
  o_{jn}.context.statement \not = \emptyset \land o_{jn} \inseq
  gl_{j})) \lor \\\t5 \langle o_{n} \rangle = \langle \rangle \rangle \\
\end{schema}

\begin{itemize}
  \item The schema $OnlyRecommendedLaunches$ updates the state defined
    by the schema $FollowedRecommendations$.
  \item $onlyRecommended'$ is the sequence of objects $o$ where $o$ is
    a sequence consisting of statements (or no statements) from the corresponding sequences
    within  $grouped_{launched}'$ where $statement.context.statement$
    exists.
  \item $onlyRecommended'$ maintains the same number and ordering of time groups
    (subsequences) as $grouped_{launched}'$ and
    $grouped_{recommended}'$.
\end{itemize}

\begin{schema}{GetCounts}
  \Delta FollowedRecommendations \\
  CountPerGroup
  \where
  cPerGroup_{launched}' = \langle c : \nat \,|\, \LET grouped_{launched}'==gl \implies
  \langle \langle gl_{i} \rangle..\langle gl_{j} \rangle \rangle @ \\\t5
  \forall \langle gl_{n} \rangle : \langle gl_{i} \rangle..\langle gl_{j} \rangle @
  \exists_1 c_{n} : \nat @ \\\t5 \IF gl_{n} = \langle \rangle \\\t6 \THEN c_{n} =
  0 \\\t6 \ELSE c_{n} = count(\langle gl_{n} \rangle)
  \rangle \\
  cPerGroup_{recommended}' = \langle c : \nat \,|\, \LET grouped_{recommended}'==gr \implies
  \langle \langle gr_{i} \rangle..\langle gr_{j} \rangle \rangle @ \\\t6
  \forall \langle gr_{n} \rangle : \langle gr_{i} \rangle..\langle gr_{j} \rangle @
  \exists_1 c_{n} : \nat @ \\\t6 \IF gr_{n} = \langle \rangle \\\t7 \THEN c_{n} =
  0 \\\t7 \ELSE c_{n} = count(\langle gr_{n} \rangle)
  \rangle \\
  cPerGroup_{followed}' = \langle c : \nat \,|\, \LET
  onlyRecommended'==or \implies \langle \langle or_{i} \rangle..\langle or_{j} \rangle \rangle @ \\\t5
  \forall \langle or_{n} \rangle : \langle or_{i} \rangle..\langle or_{j} \rangle @
  \exists_1 c_{n} : \nat @ \\\t5 \IF or_{n} = \langle \rangle \\\t6 \THEN c_{n} =
  0 \\\t6 \ELSE c_{n} = count(\langle or_{n} \rangle) \rangle \\
\end{schema}

\begin{itemize}
  \item The schema $GetCounts$ updates the state defined by the schema
    $FollowedRecommednations$.
  \item $cPerGroup_{launched}'$ is a sequence of numbers $c$ where
    each $c$ is either 0 or the result of passing the current
    subsequence of $grouped_{launched}'$ ($gl_{n}$) to the function
    $count$.
  \item  $cPerGroup_{recommended}'$ is a sequence of numbers $c$ where
    each $c$ is either 0 or the result of passing the current
    subsequence of $grouped_{recommended}'$ ($gr_{n}$) to the function
    $count$.
  \item  $cPerGroup_{followed}'$ is a sequence of numbers $c$ where
    each $c$ is either 0 or the result of passing the current
    subsequence of $onlyRecommended'$ ($or_{n}$) to the function
    $count$.
\end{itemize}

\begin{schema}{CombineSequences}
  \Delta FollowedRecommendations
  \where
  combined' = \langle c : (tr_{start}', tr_{end}', N_{launched}',
  N_{recommended}', N_{followed}', P_{followed}', P_{dueto}') \,|\, \\\t3
  \LET grouped_{launched}'==gl \implies \langle \langle gl_{i}
  \rangle..\langle gl_{n} \rangle .. \langle gl_{j} \rangle \rangle \\\t3\:\:\:\:\:\:\:
  cPerGroup_{launched}'==cl \implies \langle cl_{i}..cl_{n}..cl_{j} \rangle \\\t3\:\:\:\:\:\:\:
  cPerGroup_{recommended}'==cr \implies \langle cr_{i}..cr_{n}..cr_{j} \rangle \\\t3\:\:\:\:\:\:\:
  cPerGroup_{followed}'==cf \implies \langle cf_{i}..cf_{n}..cf_{j} \rangle \\\t3
  @ \forall \langle gl_{n} \rangle : \langle gl_{i} \rangle..\langle
  gl_{j} \rangle @ i \leq n \leq j @
  \\\t3 \exists_1 \, c_{n} : (tr_{startn}, tr_{endn}, N_{launchedn},
  N_{recommendedn}, N_{followedn}, P_{followedn}, P_{dueton}) @
  \\\t3 tr_{startn} = (head~gl_{n}).timestamp
  \\\t3 tr_{endn} = (last~gl_{n}).timestamp
  \\\t3 N_{launchedn} = cl_{n}
  \\\t3 N_{recommendedn} = cr_{n}
  \\\t3 N_{followedn} = cf_{n}
  \\\t3 P_{followedn} = cf_{n} \div cr_{n}
  \\\t3 P_{dueton} = cf_{n} \div cl_{n} \rangle
\end{schema}

\begin{itemize}
  \item The schema $CombineSequences$ changes the state defined by the
    schema $FollowedRecommendations$.
  \item $combined'$ is a sequence of objects $c$ where each $c$ is an
    ordered pair of $tr_{start}', tr_{end}', N_{launched}',
    N_{recommended}', N_{followed}', P_{followed}', P_{dueto}'$.
  \item for each $c_{n}$:
    \begin{itemize}
    \item  $tr_{start}' \bind tr_{startn}$ which is equal to the
      timestamp for the first statement within $gl_{n}$
    \item $tr_{end}' \bind tr_{endn}$ which is equal to the timestamp
      for the last statement within $gl_{n}$.
    \item $N_{launched}' \bind N_{launchedn}$ which is equal to the
      current count of launched statements within the nth time
      grouping aka $cl_{n}$.
    \item $N_{recommended}' \bind N_{recommendedn}$ which is equal to
      the current count of recommended statements within the nth time
      grouping aka $cr_{n}$.
    \item $N_{followed}' \bind N_{followedn}$ which is equal to the
      current count of recommended statements within the nth time
      grouping aka $cf_{n}$.
    \item $P_{followed}' \bind P_{followedn}$ which is equal to the
      result of dividing $cf_{n}$ by $cr_{n}$.
    \item $P_{dueto}' \bind P_{dueton}$ which is equal to the result
      of dividing $cf_{n}$ by $cl_{n}$.
    \end{itemize}
  \end{itemize}

\subsubsection{Sequence of Operations}
  $ProcessFollowedRecommendations \defs \\ OrderStatements \semi
  GroupByTime \semi OnlyRecommendedLaunches \semi \\ GetCounts \semi
  CombineSequences$
\begin{itemize}
\item The schema $ProcessFollowedRecommendations$ defines the order of
  operations for the steps within the $FollowedRecommendations$ algorithm.
\end{itemize}

\subsubsection{Return}

\begin{schema}{ReturnFollowedRecommendations}
  \Xi FollowedRecommendations \\
  ProcessFollowedRecommendations \\
  combined! : \seq
  \where
  combined! = combined'
\end{schema}

\begin{itemize}
  \item The schema $ReturnFollowedRecommendations$ describes the return
    value of the system defined by the schema
    $FollowedRecommendations$
  \item The return value $combined!$ is the variable $combined'$
    defined wtihin the schema $CombineSequences$
\end{itemize}

\subsection{Pseudocode}

\begin{algorithm}[H]
  \SetAlgoLined
  \KwIn{$S_{recommended}$, $S_{launched}$ $timeUnit$}
  \KwResult{$combined'$}
  $ordered_{L}' \leftarrow orderByTimestamp(S_{launched})$; \\
  $ordered_{R}' \leftarrow orderByTimestamp(S_{recommended})$; \\
  $t_{start}' \leftarrow convert((head~ordered_{L}').timestamp)$;\\
  $grouped_{launched}' \leftarrow groupByTimeUnit(ordered_{L}',
   t_{start}', 0, timeUnit)$;\\
  $grouped_{recommended}' \leftarrow groupByTimeUnit(ordered_{R}',
   t_{start}', 0, timeUnit)$;\\
  $grouped_{followed} \leftarrow []$;\\
   \ForEach{$G \inseq grouped_{launched}'$}
   {$curGrouping \leftarrow []$;\\
     \ForEach{$G_{n} \inseq G$}
     {\eIf{$G_{n}.context.statement \not = nil$}
       {\bf do \\
         $curGrouping' \leftarrow curGrouping \cat G_{n}$;\\
         recur curGrouping'}
       {\bf recur curGrouping'}}
   {\bf $grouped_{followed}' \leftarrow grouped_{followed} \cat
   curGrouping'$;\\
   recur $grouped_{followed}'$}}
   $C_{launched} \leftarrow {\bf map \, count() \,
     grouped_{launched}'}$;\\
   $C_{recommended} \leftarrow {\bf map \, count() \,
     grouped_{recommended}'}$;\\
   $C_{followed} \leftarrow {\bf map \, count() \,
     grouped_{followed}'}$;\\
   $combined \leftarrow []$;\\
   \For{$i\gets0$ \KwTo $count(c_{launched})$ \KwBy $1$}{
     $tr_{starti} \leftarrow (first(nth(grouped_{launched}',
     i))).timestamp$;\\
     $tr_{endi} \leftarrow (last(nth(grouped_{launched}',
     i))).timestamp$;\\
     $N_{Li} \leftarrow nth(C_{launched}, i)$;\\
     $N_{Ri} \leftarrow nth(C_{recommended}, i)$;\\
     $N_{Fi} \leftarrow nth(C_{followed}, i)$; \\
     $P_{Fi} \leftarrow N_{fi} \div N_{Ri}$;\\
     $P_{duetoi} \leftarrow N_{fi} \div N_{Li}$;\\
     $subVec_{i} \leftarrow [tr_{starti}, tr_{endi}, N_{Li}, N_{Ri},
     N_{Fi}, P_{Fi}, P_{duetoi}]$;\\
     $combined' \leftarrow combined \cat subVec_{i}$}
   \Return $combined'$
  \caption{Followed Recommendations}
\end{algorithm}
\begin{itemize}
\item {\bf map count() $grouped_{...}'$} means apply the function {\bf
  count()} to every sequence within the sequence $grouped_{...}$ and
put all results into a single array.
\end{itemize}

\subsection{JSON Schema}
\begin{lstlisting}[]
{"type":"array",
   "items":{"type":"array",
      "items":[{"type":"string"}, {"type":"string"},
               {"type":"number"}, {"type":"number"},
               {"type":"number"}, {"type":"number"},
               {"type":"number"}]}}
\end{lstlisting}
\subsection{Visualization Description}

The {\bf Followed Recommendations} visualization can be a bar chart
where the domain is time ranges and the range is a number
representing the total count of statements recorded. For each time
range, there will be three groups: 1) the number of launched
statements 2) the number of recommended statements 3) the number of
launches which are due to recommendations. Above each grouping or on
hover, summary statistics can be desplayed which describe the
percentage of launches due to recommendations and the percentage of
recommendations which were followed.

\subsection{Visualization prototype}

\begin{tikzpicture}
  \begin{axis}[
    ybar,
    enlargelimits=0.15,
    legend style={at={(0.5,-0.15)},
      anchor=north,legend columns=-1},
    ylabel={\# statements},
    symbolic x coords={8/13-8/14,8/14-8/15,8/15-8/16},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},]
    \addplot coordinates {(8/13-8/14,25) (8/14-8/15,27) (8/15-8/16,26)};
    \addplot coordinates {(8/13-8/14,16) (8/14-8/15,18) (8/15-8/16,17)};
    \addplot coordinates {(8/13-8/14,9) (8/14-8/15,7) (8/15-8/16,13)};
    \legend{Launches,Recommendations,Follows}
  \end{axis}
\end{tikzpicture}
\begin{itemize}
\item The percentages described in section 5.9 are not displayed here.
\end{itemize}

\subsection{Prototype Improvement Suggestions}
Additional features may be implemented on top of this base
specification but they would require adding aditional values to each
subarray returned by the algorithm. These additional values can be
retrieved via (1) performing metadata lookup within or independently
of the algorithm (2) by utilizing additional xAPI statement paramters
and/or (3) by performing additional computations. The following
examples assume the metadata is contained within each statement
available to the algorithm.

\begin{itemize}
\item populate a tooltip with the most popular launched, recommended
  and followed activity.
\item populate a tooltip with the number of actors associated with the
  launches and follows.
\item populate a tooltip with the actor who most often and the actor
  who lease often follows recommendations.
\end{itemize}

\end{document}
