\documentclass{article}

%%% Package imports
% cross .tex document references,
% sources: (https://texfaq.org/FAQ-extref), (http://www-users.york.ac.uk/~pjh503/LaTeX/ref.html)
\usepackage{xr}
\usepackage{xr-hyper}

\usepackage{graphicx}
\graphicspath{{../../resources/plots/} {../resources/plots/} {../../resources/figures/} {../resources/figures/}}
\usepackage{placeins}
\usepackage{subfiles}
\usepackage[utf8]{inputenc}
\usepackage[ruled]{algorithm2e}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
}
\usepackage{amsmath}
\usepackage{zed-csp}
\usepackage{breqn}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{dateplot}
\usepgfplotslibrary{polar}
\usetikzlibrary{pgfplots.dateplot}
\usetikzlibrary{pgfplots.patchplots}
\usetikzlibrary{patterns}

\usepackage{floatrow}

\usepackage{calc}
%%% \makeatletter\amparswitchfalse\makeatother\
\DeclareMarginSet{hangleft}{\setfloatmargins
{\hskip-\marginparwidth\hskip-\marginparsep}{\hfil}}
\floatsetup[widefigure]{margins=hangleft}
%%% ^ Figure formatting within Appendex A

\lstset{literate = {-}{-}1} % get dashs to show up

\pgfplotsset{compat=1.15}

\SetKw{KwBy}{by}

\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage}

\title{Data Analytics and Visualization Environment for xAPI and the Total Learning Architecture: DAVE Learning Analytics Algorithms}
\author{Yet Analytics}

\begin{document}

\begin{titlepage}
  \maketitle
\end{titlepage}

\section*{Introduction}

This report introduces a language for defining the functionality of learning analytics algorithms in terms of
\textbf{Operations}, \textbf{Primitives} and \textbf{Algorithms} which will be used to define Algorithms corresponding to
an inital set of learning analytics questions. Additional questions may be added to this set in the future.
This document will be updated to include additional Operations, Primitives and Algorithms as they are defined by the Author(s)
of this report or by members of the Open Source Community. Updates may also address refinement of existing definitions,
thus this document is subject to continious change but those which are significant will be documented within the DAVE change log.
Any changes made to this report or to the \href{https://github.com/yetanalytics/dave}{DAVE github repository} should follow
the conventions established in the \href{https://github.com/yetanalytics/dave/wiki/Contributing}{Contributing Wiki Page}.
The formal definitions in this document are optimized for understandability and conceptual presentation meaning they are not
presented as, or intended to be, the most computationaly effecient definition possible. The formal definitions are intended
to serve as referential documentation of methedologies and programatic strategies for handling the processing of xAPI data.
$\\\\$
The structure of this documents is as follows:
\begin{enumerate}
\item An Introduction to Z notation and its usage in this document
\item A formal specification for xAPI written in Z % FIXME: review + refactor as necessary
\item Terminology: Operations, Primitives and Algorithms
\item What is an Operation
\item What is a Primitive
\item What is an Algorithm
\item Foundational Operations
\item Common Primitives
\item Example Algorithm
  \begin{enumerate}
  \item Init
  \item Relevant?
  \item Accept?
  \item Step
  \item Result
  \end{enumerate}
\end{enumerate}

% FIXME: direct section references using cross-refering mechanism
% - \label + \externaldocument + \nameref, \ref, pageref

\subfile{z/introduction.tex}
\subfile{z/xapi.tex}
\subfile{algorithms/introduction.tex}

\section{Foundational Operations}
The Operations in this section use the Operations pulled from the Z Reference Manual (section 1,4) within their own definitions.
They are defined as Operations opposed to Primitives because they represent core functionality needed in the context
of processing xAPI data given the definition of an Algorithm above. As such, these Operations are added to the global
dictionary of symbols usable, without a direct reference to the components schema, within the definition of
Operations and Primitives throughout the rest of this document. In general, Operations are intended to be
simple, and should not contain any recursive calls. They are building blocks which are used across Primitives
of varying functionality. When defining an Operation not already in the set of Foundational Operations defined here, its schema
MUST be referenced at the top of all Schemas which utilize the new Operation.

\subsection{Collections}
Operations which expect a Collection $X = \langle x_{i}..x_{n}..x_{j} \rangle$
\subfile{operations/collections/array?.tex}
\subfile{operations/collections/append.tex}
\subfile{operations/collections/remove.tex}
\subfile{operations/collections/atIndex.tex}
\subfile{operations/collections/update.tex}

\subsection{Key Value Pairs}
Operations which expect a Map $M = \ldata k_{i}v_{k_{i}}..k_{n}v_{k_{n}}..k_{j}v_{k_{j}} \rdata$

\subfile{operations/kv/map?.tex}
\subfile{operations/kv/associate.tex}
\subfile{operations/kv/dissociate.tex}
\subfile{operations/kv/atKey.tex}

\subsection{Utility}
Operations which are usefull in many Statement processing contexts.
\subfile{operations/util/map.tex}
\subfile{operations/util/isoToUnix.tex}
\subfile{operations/util/timeUnitToNumberOfSeconds.tex}
\subfile{operations/util/rateOf.tex}
% FIXME: Define atJSONPath as Operation (special case, hence utility categorization).
% - Reference another definition of atJsonPath by URL within copy and establish likeness to getIn
% - in the case of JSONPath, the arg $path?$ is a domain specific structured string but that language is defined outside of this document
% - further details are outside the scope of this document
% - update file path, prev version = \subfile{primitives/atJsonPath.tex}

\section{Common Primitives}
% Low Priority:
% TODO: update file name?
% TODO: split into multiple files?
% TODO: clean~\_ primitive - removes all mappings where V \mapsto \emptyset
\subfile{primitives/walk.tex}

% New section handled within rateOfCompletions
\subfile{algorithm_definitions/rateOfCompletions.tex}

% TODO: update legacy versions to Z
%\subfile{algorithm_definitions/timelineLearnerSuccess.tex}
%\subfile{algorithm_definitions/mostDifficultAssessmentQuestions.tex}
%\subfile{algorithm_definitions/followedRecommendations.tex}

% TODO: remove references to legacy versions of Algorithms
%\subfile{algorithms/rate_of_completions.tex}
%\subfile{algorithms/timeline_learner_success.tex}
%\subfile{algorithms/most_difficult_assessment_questions.tex}
%\subfile{algorithms/followed_recommendations.tex}

\subfile{appendices/a.tex}

\end{document}
